### 2017-04-16 proposal.tex
Boilerplate

% Template from Usenix
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\begin{document}

% Don't output date
\date{}

% Title
\title{\Large \textbf{
(TODO: Project Name) - Project Proposal } \\ \vspace{0.025 in} \large \normalfont
CSE 227: Computer Security - Spring 2017 \\ \textit{
University of California San Diego
}}

% Authors
\author{
{\rm Chen Lai}\\
\normalfont{\texttt{chl588@ucsd.edu}}
\and
{\rm Zhongrong Jian}\\
\normalfont{\texttt{zhjian@ucsd.edu}}
\and
{\rm Juan Sidrach}\\
\normalfont{\texttt{jsidrach@ucsd.edu}}
}

\maketitle

\section{Introduction}
TODO

\section{Section}
TODO~\cite{ipv4sta}

{\footnotesize \bibliographystyle{acm}
\bibliography{bibliography}}

\end{document}